\Title

### 2011-04-18 masters_thesis/thesis/paper.tex
Added a thesis version of my thesis

\documentclass[smallcondensed]{svjour3}

\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{psfrag}
\usepackage{natbib}

@@

\title{A Speculative Approach to Parallelization in Particle Swarm
Optimization}

\date{Received: date / Accepted: date}

\author{Matthew Gardner, Andrew McNabb, and Kevin Seppi}

\institute{M. Gardner \at
Brigham Young University.
3361 TMCB, Provo, UT 84604
Tel.: 801-422-8717
Fax: 801-422-0169
\email{mjg82@byu.edu}
\and
A. McNabb \at
Brigham Young University.
3361 TMCB, Provo, UT 84604
\email{a@cs.byu.edu}
\and
K. Seppi \at
Brigham Young University.
3361 TMCB, Provo, UT 84604
\email{k@cs.byu.edu}
}

\begin{document}
\maketitle

\begin{abstract}

Particle swarm optimization (PSO) has previously been parallelized primarily by
distributing the computation corresponding to particles across multiple
processors.  In these approaches, adding processors can only increase the swarm
size.  However, in many cases this is not efficient when scaled to very large
swarm sizes (on very large clusters).  Current methods cannot answer well the
question, ``How can 1000 processors be fully utilized when 50 or 100 particles
is the most efficient swarm size?''  In this paper we attempt to answer that
question with a speculative approach to the parallelization of PSO that we
refer to as SEPSO.

@@

However, given this new parallelization model we can relax the requirement of
exactly reproducing PSO in an attempt to produce better results.  We present
several such relaxations, including keeping the best speculative position
evaluated instead of the one corresponding to the standard behavior of PSO, and
speculating several iterations ahead instead of just one.  We show that these
methods dramatically improve the performance of parallel PSO in many cases.

\keywords{Parallel algorithms \and Optimization methods \and Particle swarm
optimization \and Speculative Decomposition}

\end{abstract}

\section{Introduction}
\label{sec:intro}

@@

\section{Speculative Evaluation in PSO}
\label{sec:sepso}

@@

\subsection{Refactoring the PSO Equations}
\label{sec:proof}

@@

\subsection{Topology in Speculative Evaluation}
\label{sec:topology}

@@

\section{Relaxing the Requirements}
\label{sec:relax}

@@

\subsection{Pick the Best Child}
\label{sec:pickbest}

@@

\subsection{Pruning the Speculative Evaluations}
\label{sec:pruning}

@@

\subsection{Branch Statistics}
\label{sec:statistics}

@@

\subsection{Recovering from Pruning Too Much}
\label{sec:wrong}

@@

\subsection{More than one iteration ahead}
\label{sec:manyiters}

@@

\section{Experimental Setup}
\label{sec:setup}

@@

\subsection{Function Evaluations vs. Time Steps}

@@

\subsection{Functions Used}

@@

\subsection{Parallelization Techniques Compared}

@@

\subsection{Topologies}

@@

\section{Results}
\label{sec:results}

@@

\subsection{20 Dimensions}
\label{sec:20dims}

\subsubsection{Sphere}
\label{sec:sphere20dims}

@@

\subsubsection{Schwefel 2.21}

@@

\subsubsection{Rastrigin}

@@

\subsubsection{Griewank}
\label{sec:griewank-20}

@@

\subsubsection{Bohachevsky}

@@

\subsection{50 Dimensions}
\label{sec:50dims}

@@

\subsection{500 Dimensions}
\label{sec:500dims}

@@

\subsection{Model Fitting}
\label{sec:model}

@@

\subsection{Bare Bones PSO}
\label{sec:bare-results}

In order to demonstrate that our methods are applicable across PSO variants, we
ran experiments with Bare Bones PSO.  For this simple experiment we use the
Sphere function with 20 dimensions, taking the best Standard algorithm and the
best speculative algorithm on that function, which in this case is Standard
Complete and Pick Best Pruned Complete, as shown in \sec{sphere20dims}.
\fig{bare-bones} shows the results of running each of those algorithms with the
Bare Bones PSO equations.  We see a decrease in time to completion between
Standard and speculative parallelizations of 40\%.  While this is not quite as
large a speed up as the 60\% seen with constricted PSO, it is still a
significant decrease in time.  The key point of this experiment is that our
methods still show improvements over Standard parallelizations in PSO variants
other than the standard constricted PSO.  Our intent is not to give an
exhaustive listing of PSO variants and compare parallelization strategies with
each one; instead we offer evidence supporting our claim that speculative
evaluation should see similar increases in performance across most or all PSO
variants.

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{bare_bones}
  \caption{Results for optimizing the 20-dimensional Sphere function with Bare
  Bones PSO.  We use 240 processors for each method, so each time step
  corresponds to 240 function evaluations.}
  \label{fig:bare-bones}
\end{figure}

\section{Conclusions}
\label{sec:conclusion}

@@

\section{Future Work}
\label{sec:future}

@@

\section*{Appendix A---Implementing Speculative Evaluation}
\label{sec:implementation}

@@

\subsection*{Terminology}

@@

\subsection*{Centralized Algorithms}

@@

\subsection*{Distributed Algorithms}

@@

\subsection*{Dynamic Topologies}

@@

\section*{Appendix B---Alternate form of message passing}

@@

\bibliographystyle{apa-good}
\bibliography{%
../../../bib/aml/bib,%
../../../bib/nfl/bib,%
../../../bib/optimization/bib,%
../../../bib/parallel/bib,%
../../../bib/pso/general/bib,%
../../../bib/pso/parallel/bib,%
../../../bib/pso/fips/bib,%
../../../bib/pso/topology/bib,%
../../../bib/sim-anneal/parallel/bib,%
../../../bib/ga/bib}

>>>>

\documentclass[ms,electronic,twosidetoc,letterpaper,chaptercenter,parttop,equalmargins]{byumsphd}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thesis formating stuff
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\Title}{A Speculative Approach to Parallelization\\
in Particle Swarm Optimization}
\newcommand{\Author}{Matthew Gardner}
\newcommand{\GraduationMonth}{August}
\newcommand{\GraduationYear}{2011}

\usepackage[
    bookmarks=true,
    bookmarksnumbered=true,
    breaklinks=false,
    raiselinks=true,
    pdfborder={0 0 0},
    colorlinks=false,
    plainpages=false,
    ]{hyperref}
\usepackage[all]{hypcap}
\usepackage{natbib}
\hypersetup{%
    pdftitle=\Title,%
    pdfauthor=\Author,%
    pdfsubject={PhD Dissertation, BYU CS Department: %
                Degree Granted \GraduationMonth~\GraduationYear, Document Created \today},%
    pdfkeywords={BYU, thesis, dissertation, LaTeX},%
}
\newcommand{\ItemSep}{\itemsep 0pt}
\let\oldenum=\enumerate
\renewcommand{\enumerate}{\oldenum \ItemSep}
\let\olditem=\itemize
\renewcommand{\itemize}{\olditem \ItemSep}
\let\olddesc=\description
\renewcommand{\description}{\olddesc \ItemSep}

% Important settings for the byumsphd class.
\title{\Title}
\author{\Author}

\committeechair{Kevin~Seppi}
\committeemembera{Dan~Ventura}
\committeememberb{David~Embley}

\monthgraduated{\GraduationMonth}
\yeargraduated{\GraduationYear}
\yearcopyrighted{\GraduationYear}

\department{Computer~Science}
\graduatecoordinator{Kent~E.~Seamons}
%\collegedean{Thomas~W.~Sederberg}
%\collegedeantitle{Associate~Dean}

% Remove all widows an orphans.  This is not normally recommended, but in a
% paper dissertation there is no reasonable way around it; you can't exactly
% rewrite already-published content to fix the problem.
\clubpenalty 10000
\widowpenalty 10000

% Allow pages to have extra blank space at the bottom in order to accommodate
% removal of widows and orphans.
\raggedbottom

% Produce nicely formatted paragraphs. There is nothing additional to do.  In
% case you get some problems, surround your text with
% \begin{sloppy} ... \end{sloppy}. If that does not work, try
% \microtypesetup{protrusion=false} ... \microtypesetup{protrusion=true}
\usepackage{microtype}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For the content of the document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{psfrag}
\usepackage{natbib}

@@

\documentabstract{%

Particle swarm optimization (PSO) has previously been parallelized primarily by
distributing the computation corresponding to particles across multiple
processors.  In this thesis we present a speculative approach to the
parallelization of PSO that we refer to as SEPSO.

@@

However, given this new parallelization model we can relax the requirement of
exactly reproducing PSO in an attempt to produce better results.  We present
several such relaxations, including keeping the best speculative position
evaluated instead of the one corresponding to the standard behavior of PSO, and
speculating several iterations ahead instead of just one.  We show that these
methods dramatically improve the performance of parallel PSO in many cases.
}
\documentkeywords{%
Parallel algorithms, optimization methods, particle swarm optimization,
peculative decomposition
}

\begin{document}

% Produce the preamble
\microtypesetup{protrusion=false}
\maketitle
\microtypesetup{protrusion=true}

\chapter{Introduction}
\label{sec:intro}

@@

\chapter{Speculative Evaluation in PSO}
\label{sec:sepso}

@@

\section{Refactoring the PSO Equations}
\label{sec:proof}

@@

\section{Topology in Speculative Evaluation}
\label{sec:topology}

@@

\chapter{Relaxing the Requirements}
\label{sec:relax}

@@

\section{Pick the Best Child}
\label{sec:pickbest}

@@

\section{Pruning the Speculative Evaluations}
\label{sec:pruning}

@@

\section{Branch Statistics}
\label{sec:statistics}

@@

\section{Recovering from Pruning Too Much}
\label{sec:wrong}

@@

\section{More than one iteration ahead}
\label{sec:manyiters}

@@

\chapter{Experimental Setup}
\label{sec:setup}

@@

\section{Function Evaluations vs. Time Steps}

@@

\section{Functions Used}

@@

\section{Parallelization Techniques Compared}

@@

\section{Topologies}

@@

\chapter{Results}
\label{sec:results}

@@

\section{20 Dimensions}
\label{sec:20dims}

\subsection{Sphere}

@@

\subsection{Schwefel 2.21}

@@

\subsection{Rastrigin}

@@

\subsection{Griewank}
\label{sec:griewank-20}

@@

\subsection{Bohachevsky}

@@

\section{50 Dimensions}
\label{sec:50dims}

@@

\section{500 Dimensions}
\label{sec:500dims}

@@

\section{Model Fitting}
\label{sec:model}

@@

\chapter{Conclusions}
\label{sec:conclusion}

@@

\chapter{Future Work}
\label{sec:future}

@@

\appendix
\chapter{Implementing Speculative Evaluation}
\label{sec:implementation}

@@

\section{Terminology}

@@

\section{Centralized Algorithms}

@@

\section{Distributed Algorithms}

@@

\section{Dynamic Topologies}

@@

\chapter{Alternate form of message passing}

@@

\bibliographystyle{plainnat}
\bibliography{%
../../../bib/aml/bib,%
../../../bib/nfl/bib,%
../../../bib/optimization/bib,%
../../../bib/parallel/bib,%
../../../bib/pso/general/bib,%
../../../bib/pso/parallel/bib,%
../../../bib/pso/fips/bib,%
../../../bib/pso/topology/bib,%
../../../bib/sim-anneal/parallel/bib,%
../../../bib/ga/bib}
